### 2025-03-17 QM/paper.tex
add a paper draft

\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{authblk}
\usepackage{float}

\title{A Historical Analysis of Quark Matter Conferences: \\
Trends, Participation, and Topic Evolution}

\author[1]{D.J~Kim}
\author[1]{C.~Sporleder}
\author[1]{P.~Runko}
\author[1]{T.~Lappeteläinen}
\affil[1]{Department of Physics, University of jyväskylä, Finland}

\begin{document}

\maketitle

\begin{abstract}
This paper presents a comprehensive analysis of Quark Matter conferences from 2011 to 2025, examining the evolution of research topics, institutional participation, and geographical distribution of contributions. By analyzing presentation data from the conference proceedings, we identify key trends in heavy-ion physics research, changes in participation patterns, and shifts in research focus over time. Our findings reveal how the field has evolved and provide insights that can help improve future conference planning and increase diversity of participation. This historical perspective offers valuable information for the heavy-ion physics community and conference organizers seeking to enhance the inclusivity and representation at these important scientific gatherings.
\end{abstract}

\section{Introduction}

The Quark Matter conference series represents the premier international meeting dedicated to ultra-relativistic heavy-ion collisions and the study of nuclear matter under extreme conditions. Since its inception, these conferences have served as crucial platforms for presenting breakthrough research, fostering collaborations, and shaping the direction of the field.

Understanding the historical patterns and evolution of these conferences provides valuable insights into:
\begin{itemize}
    \item The shifting landscape of research topics and methodologies
    \item Geographical and institutional representation in the field
    \item Opportunities to improve diversity and inclusivity in scientific participation
    \item The impact of major experimental programs and theoretical advances
\end{itemize}

In this paper, we analyze data from Quark Matter conferences spanning from 2011 to 2025, extracting patterns from presentation titles, speaker affiliations, and presentation categories. Our goal is to provide a quantitative assessment of how the field has evolved and identify areas where conference organization might be improved to better serve the scientific community.

\section{Data and Methods}

\subsection{Data Collection}

We collected data from the official Indico pages of Quark Matter conferences from 2011 to 2025. For each conference, we extracted:
\begin{itemize}
    \item Presentation titles, authors, and affiliations
    \item Presentation categories (plenary, parallel, poster, flash)
    \item Location and year information
\end{itemize}

A custom Python script was developed to fetch this data using the Indico API. The script processes the contribution data, categorizes talks, and extracts relevant information about speakers and their affiliations.

\subsection{Data Processing}

Our analysis pipeline included several key processing steps:
\begin{itemize}
    \item Classification of presentations into plenary, parallel, poster, and flash categories
    \item Extraction of speaker information including names, institutes, and countries
    \item Text processing of presentation titles to identify key research topics
    \item Aggregation of statistics by conference, institute, and country
    \item Implementation of manual corrections for missing or incorrect data
\end{itemize}

To identify research trends, we performed keyword extraction from presentation titles after removing common stop words and non-meaningful terms. This allowed us to track the evolution of research focus across different conferences.

Despite our extensive efforts to map institutes to countries, a small number of institute entries (14 in total) could not be reliably mapped to countries. These include single-letter entries (B, D, F, I, L, N, R, U, V, W, l), generic labels (PhD student), and undefined entries (Unknown, Unknown-Unknown-Unknown). These entries likely represent data entry errors or incomplete information in the original conference records rather than actual institutes. In our analysis, these were categorized as "Unknown" country to maintain data integrity.

\section{Results}

\subsection{Conference Venues and Geographical Scope}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/conference_venues.pdf}
\caption{Geographical distribution of Quark Matter conference venues from 2011 to 2025. The timeline displays each conference location, color-coded by continent. This visualization illustrates the global reach of the conference series and regional patterns in host selection over time.}
\label{fig:venues}
\end{figure}

Figure~\ref{fig:venues} presents the geographical distribution of Quark Matter conference venues over time. This visualization offers important context for understanding the evolution of the conference series and its global outreach efforts:

The venue selection shows a deliberate effort to rotate the conference across different global regions, with representation from North America, Europe, and Asia throughout the analyzed period. This rotation policy helps ensure that the conference remains accessible to researchers from different parts of the world over time, even if individual conferences may have geographical attendance biases.

There appears to be a pattern of alternating between continents, particularly between Europe and North America, with Asian venues interspersed at less regular intervals. This pattern reflects both the historical centers of heavy-ion physics research and efforts to expand the global footprint of the conference.

The frequency of venues in different regions roughly corresponds to the size of the heavy-ion physics community in those areas, with Europe and North America hosting most frequently, followed by Asia. However, there are notable gaps in geographical coverage, particularly from South America, Africa, and parts of Asia beyond East Asia.

The choice of venue has important implications for participation patterns, as shown in our subsequent analyses of speaker demographics. Conference attendance is typically higher from the host country and region, affecting both the volume and diversity of submissions from different geographical areas.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/QM_talk_statistics.pdf}
\caption{Statistical overview of presentations at Quark Matter conferences from 2011 to 2025. The chart shows the distribution of different presentation types (plenary, parallel, poster) across conference years, illustrating changes in conference structure and overall participation volume over time.}
\label{fig:talk_statistics}
\end{figure}

Figure~\ref{fig:talk_statistics} provides a statistical overview of presentations at Quark Matter conferences over the analyzed period. Several notable patterns emerge from this data:

The total number of presentations varies significantly between conferences, reflecting both changes in the size of the heavy-ion physics community and practical constraints of different venues. This variation affects the competitiveness of selection processes, particularly for high-visibility plenary and parallel talks.

The ratio between different presentation types (plenary, parallel, poster) has evolved over time, with recent conferences generally allocating a larger proportion of slots to poster presentations. This shift may reflect efforts to accommodate growing participation while maintaining selectivity for oral presentations.

The introduction of flash talk sessions in more recent conferences represents an innovation in conference format, providing brief speaking opportunities for a larger number of participants while maintaining the benefits of poster-based detailed discussions.

These structural changes in conference organization have important implications for participation patterns and visibility distribution across different research groups and geographical regions, as explored in our subsequent analyses.

\subsection{Research Topic Evolution}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/top_keywords.pdf}
\caption{Evolution of top research topics in Quark Matter conferences from 2011 to 2025. Each cluster of bars represents a conference year, with the seven most frequent keywords from presentation titles shown with their occurrence counts. This visualization reveals shifts in the field's focus over time, highlighting emerging and declining research areas.}
\label{fig:keywords}
\end{figure}

Figure~\ref{fig:keywords} presents the evolution of research topics across Quark Matter conferences from 2011 to 2025, as revealed by keyword frequency analysis of presentation titles. Several notable trends emerge from this visualization:

The keyword "QGP" (Quark-Gluon Plasma) consistently appears among the top terms across most conference years, reflecting the central focus of the field. However, its relative prominence varies over time, suggesting shifts in how directly presentations address the QGP versus specific properties or phenomena within it.

We observe a clear evolution in experimental systems of interest. Early conferences show significant focus on "LHC" and "RHIC" accelerator facilities, while later years show more diversified keywords related to specific measurements and phenomena, indicating a maturation of the field from facility-focused to phenomenon-focused research.

Certain physics concepts show interesting temporal patterns. Terms related to "flow" and "collectivity" are prominent in middle years, corresponding to the period of intense investigation into hydrodynamic behavior in heavy-ion collisions. Later conferences show increased occurrence of terms related to "small systems" and "fluctuations," reflecting the field's expanding interest in these areas.

Theoretical concepts also show clear evolution, with early emphasis on "model" and "simulation" gradually giving way to more specific theoretical approaches and phenomena, suggesting increasing theoretical sophistication and specialization.

This keyword analysis provides a data-driven view of how research priorities in the heavy-ion physics community have evolved over the past decade and points to potential future directions as indicated by emerging keywords in recent conferences.

\subsection{Geographical Distribution of Contributions}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/plenary_talks_by_country.pdf}
\caption{Distribution of plenary talks by country across Quark Matter conferences from 2011 to 2025. The stacked bars show the number of plenary presentations given by speakers from each of the top 10 countries. This visualization reveals the geographical concentration of high-visibility presentations and changes in international representation over time.}
\label{fig:country_plenary}
\end{figure}

Figure~\ref{fig:country_plenary} displays the distribution of plenary talks by country across Quark Matter conferences. Plenary talks represent the highest visibility presentations at these conferences and are typically allocated to highlight significant advances in the field. Several important patterns are evident:

There is a consistent dominance by a small number of countries, particularly the United States, which maintains a substantial share of plenary talks across all conference years. This reflects the significant investment in heavy-ion research infrastructure and personnel in the US, home to the RHIC facility and major ALICE, CMS, and ATLAS heavy-ion programs.

European countries collectively represent another major block, with Germany, France, and the UK consistently present. The distribution among European countries shows some variability between conferences, potentially reflecting both the location of the conference (European conferences tend to have more European speakers) and shifts in research output.

Asian representation, particularly from China, Japan, and India, shows interesting dynamics over time. We observe a general trend of increasing representation from these countries, especially in more recent conferences, reflecting growing investment in the field in these regions.

The data reveals potential geographical imbalances in high-visibility speaking opportunities. While some variation is expected due to differences in community size and research output, the persistence of these patterns may warrant attention from conference organizers interested in ensuring equitable international representation.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/parallel_talks_by_country.pdf}
\caption{Distribution of parallel talks by country across Quark Matter conferences from 2011 to 2025. The stacked bars show the number of parallel session presentations given by speakers from each of the top 10 countries. This visualization reveals broader participation patterns compared to plenary talks and illustrates how research contributions are distributed globally.}
\label{fig:country_parallel}
\end{figure}

Figure~\ref{fig:country_parallel} shows the distribution of parallel talks by country, providing insight into the broader participation patterns beyond the highly selective plenary sessions. Parallel talks represent the majority of oral presentations at Quark Matter conferences and offer a more comprehensive view of active research in the field:

The country distribution for parallel talks shows greater diversity compared to plenary talks, with more countries represented and somewhat more balanced proportions. This suggests that while plenary selections may favor established research groups from certain countries, the parallel sessions incorporate a wider geographical range of contributions.

We observe interesting temporal trends, with some countries showing increasing representation over time (e.g., China), while others maintain relatively stable participation. These trends likely reflect both changes in research output and evolving selection processes by conference organizers.

The correlation between conference location and country representation is somewhat visible in the parallel session data, with host countries typically showing increased representation in their conference years. This may reflect both practical considerations (travel funding) and deliberate efforts to showcase local research.

Despite the greater diversity, there remains a clear stratification, with a few countries consistently accounting for the majority of parallel talks. This pattern raises questions about structural factors affecting international participation, including funding disparities, language barriers, and access to research infrastructure.

\subsection{Institutional Representation}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/plenary_talks_by_institute.pdf}
\caption{Distribution of plenary talks by institute across Quark Matter conferences from 2011 to 2025. The stacked bars show the number of plenary presentations given by speakers from each of the top 15 institutes. This visualization reveals which research institutions have the most prominent representation in high-visibility speaking slots, highlighting potential concentration of influence in the field.}
\label{fig:institute_plenary}
\end{figure}

Figure~\ref{fig:institute_plenary} presents the distribution of plenary talks by research institute across Quark Matter conferences, offering insight into which institutions have the most significant presence in these high-visibility presentations:

The data reveals a strong concentration of plenary talks among a small group of elite research institutions, primarily major national laboratories and top-tier universities with strong nuclear physics programs. This concentration is even more pronounced than the country-level analysis, suggesting that institutional affiliation may be a more significant factor than nationality in plenary speaker selection.

Major experimental facilities hosts, such as Brookhaven National Laboratory (RHIC), CERN (LHC), and GSI (FAIR), are consistently well-represented, reflecting their central role in generating the data that drives the field forward. University representation is dominated by institutions with strong theoretical groups or significant roles in detector development and data analysis.

The temporal trends show some evolution in institutional representation, but the changes are relatively modest compared to the persistent dominance of established centers. New entrants to the "top institutes" list are rare across the conference series, suggesting limited mobility in the institutional hierarchy.

These patterns raise important questions about how speaking opportunities, particularly high-visibility plenary talks, are allocated within the field. While meritocratic considerations certainly play a role, the strong institutional concentration may create barriers for researchers from less prestigious institutions to gain recognition for their work.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/parallel_talks_by_institute.pdf}
\caption{Distribution of parallel talks by institute across Quark Matter conferences from 2011 to 2025. The stacked bars show the number of parallel session presentations given by speakers from each of the top 15 institutes. This visualization reveals broader institutional participation patterns compared to plenary talks and illustrates how research contributions are distributed across research centers.}
\label{fig:institute_parallel}
\end{figure}

Figure~\ref{fig:institute_parallel} shows the distribution of parallel talks by institute, providing a broader view of institutional participation beyond the highly selective plenary sessions:

Similar to the country-level analysis, we observe greater institutional diversity in parallel sessions compared to plenary talks. While the leading institutions maintain strong representation, many more institutions participate at this level, creating a longer "tail" of participation that includes smaller universities and research groups.

The data shows some interesting dynamics in institutional participation over time. Some institutions show surges in certain years, often corresponding to significant experimental results or theoretical advances from those groups. Others maintain more consistent participation levels across the conference series.

Even in this broader participation category, there remains a clear stratification, with approximately 15-20 institutions accounting for a disproportionate share of all parallel talks. This concentration, while less extreme than for plenary talks, still suggests significant disparities in research visibility.

The institutional distribution also reflects the collaborative nature of the field, with participating institutions often clustered around major experimental collaborations (ALICE, STAR, PHENIX, etc.). Institutions with prominent roles in these collaborations typically secure more speaking slots, reflecting both their scientific contributions and their structural positions within the field.

\section{Discussion}

\subsection{Diversity and Representation}

Our analysis of geographical and institutional representation at Quark Matter conferences reveals persistent patterns that merit careful consideration. The concentration of presentations, particularly high-visibility plenary talks, among a small group of countries and institutions suggests structural imbalances in how research visibility is distributed in the field.

While some degree of concentration is expected given differences in community size, research infrastructure, and historical development of the field in different regions, the persistence of these patterns across multiple conference cycles raises questions about the mechanisms of speaker selection and the potential for implicit biases in these processes.

The data suggests several potential approaches for conference organizers interested in enhancing diversity of representation:

\begin{itemize}
    \item Implementing transparent selection criteria that explicitly value geographical and institutional diversity alongside scientific excellence
    \item Developing mentoring programs or speaking opportunities specifically designed to highlight researchers from underrepresented institutions and countries
    \item Rotating conference locations more systematically to encourage broader participation and reduce travel barriers
    \item Collecting and monitoring demographic data on conference participation to track progress over time
\end{itemize}

Importantly, our findings suggest that addressing these imbalances may require attention not just to country-level representation but to institutional diversity within countries, as the institutional concentration appears even stronger than geographical concentration.

\subsection{Evolution of Research Focus}

The keyword analysis provides a fascinating window into how the scientific focus of the heavy-ion physics community has evolved over the past decade. Several key patterns emerge:

\begin{itemize}
    \item A shift from facility-focused research (emphasized by accelerator and detector keywords) toward more phenomenon-focused investigations
    \item The rise and fall of certain research themes, reflecting the community's response to new discoveries and theoretical developments
    \item Increasing specialization and technical sophistication, as indicated by the growing specificity of keywords in later conferences
    \item The emergence of new research directions, particularly investigations of collectivity in small systems and the application of machine learning techniques
\end{itemize}

These trends reflect a field in transition, moving from an early phase focused on establishing the basic properties of the quark-gluon plasma toward more detailed investigations of specific phenomena and increasingly precise measurements. The keyword dynamics also highlight the responsiveness of the field to surprising experimental results, such as the observation of collective behavior in small collision systems.

\subsection{Implications for Conference Organization}

Based on our findings, several recommendations emerge for future Quark Matter conference planning:

\begin{itemize}
    \item Implementation of more transparent and explicitly diversity-conscious selection processes for talks, particularly plenary presentations
    \item Consideration of mechanisms to provide speaking opportunities for researchers from underrepresented countries and institutions
    \item Balancing representation of established research directions with emerging topics to ensure the conference remains at the cutting edge
    \item Improving data collection on speaker demographics to better track progress on diversity goals
    \item Developing structures to highlight work from smaller or newer research groups, potentially through dedicated sessions or presentation formats
\end{itemize}

These measures would help ensure that Quark Matter conferences continue to serve as inclusive forums that represent the full breadth of the international heavy-ion physics community.

\section{Conclusion}

This historical analysis of Quark Matter conferences provides valuable insights into the evolution of heavy-ion physics as reflected in its premier conference series. By quantifying patterns of participation, geographical distribution, and research focus, we have established a factual basis for discussions about how these conferences might better serve the scientific community.

Our findings reveal both strengths and areas for improvement in the current structure of these conferences. The keyword analysis demonstrates a vibrant field addressing an evolving set of scientific questions with increasing sophistication. At the same time, the geographical and institutional analyses highlight persistent patterns of concentration that may not fully reflect the global nature of the field.

As Quark Matter conferences continue to evolve, regular analysis of participation patterns can provide valuable feedback to organizers and the community as a whole. By making these patterns visible, we hope to contribute to ongoing discussions about how to ensure these important scientific gatherings reflect the diversity and dynamism of heavy-ion physics research worldwide.

\section{Acknowledgments}

We thank the organizers of Quark Matter conferences for making the presentation data publicly available through Indico, enabling this analysis.

\bibliographystyle{unsrt}
\begin{thebibliography}{9}
\bibitem{indico} 
CERN. ``Indico - Event Management System.'' 
\texttt{https://indico.cern.ch/}

\bibitem{QM2011}
Quark Matter 2011 Conference Proceedings,
\texttt{https://indico.cern.ch/event/30248/}

% Additional references would be added here
\end{thebibliography}

\end{document}